\textbf{ Spectrophotometric parallaxes with \project{The Cygnet}:\\ Accurate distances for luminous red-giant stars }

### 2018-08-06 paper/sp.tex
linearity assumption

In that context: This model is only linear! We could have done much better with
a more flexible model. But choices, execution.

@@

>>>>

@@

As we said above, our most restrictive assumption is the linearity assumption.
We required that the parallax prediction be constructed from a linear combination
of the input data or feature vector components.
This is absurd: Stars are not linear objects!
There are two questions to ask about the linearity.
One is: Why did it work? The other is: Could we have done better?
The answer to the first is that we have taken such a small part of stellar parameter
space with our surface gravity cut, that the linearized model of stellar
spectrophotometry is not a bad approximation.
Also, because our linear model included photometry in magnitudes (that is, logarithmic)
and was exponentiated to make the parallax prediction, we knew in advance that much
of the needed flexibility (for dust attenuation, for example) would be close to
linear.
That is, it was a combination of limited model scope and some cleverness in the
feature engineering and model structure.

The answer to the second question is that absolutely we could have done better!
Since models like Gaussian processes and deep networks subsume the linear model
(at least approximately), they could have delivered similar results.
The issues with going to a much more complex model are manifold, however:
We would have had many more decisions to make and many more hyper-parameters
to set.
We would have used some (or maybe much) of the information in the data to learn
the simple fact that the problem is close to linear; that is, if you give a model
a lot of freedom or capacity, it has to use a lot of the data to learn what part
of that capacity it really needs.
A more complex model might have had edge issues: Very flexible models don't extrapolate
outside the convex hull of their training data, and our training data was probably not
balanced in terms of parallax and dust attenuation.
And finally, the complex model would have been harder to visualize and interpret;
the linear model contains essentially only derivatives in the data space, which are
relatively easy to understand.
